---
title: 'MSDS-692: Data Science Practicum I'
subtitle: 'Data Ingest From S3'
author: "Thomas Skowronek"
date: "January 27, 2018"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
  word_document: default
---


```{r setup-layout, cache=FALSE, echo=FALSE}
# Define the report layout.
library(knitr)

# Set global chunk options: images sizes for html and docx
output <- opts_knit$get("rmarkdown.pandoc.to")

if (output=="html") opts_chunk$set(fig.width=10, fig.height=6)
if (output=="docx") opts_chunk$set(fig.width=10,  fig.height=6)

# Set the printable width
options(width = 85)
```
\newline


## Setup
Set the working directory, clear all existing objects in the workspace and set the seed for reproducibility.  Lastly, load the necessary libraries.
```{r env-config, warning=FALSE, message=FALSE}
# Set the working directory
setwd("./")

# Clear all existing objects in the workspace
rm(list = ls())

# Set the seed for reproducible results
set.seed(1009)

# Load libraries
library(downloader)
```
\newline


## Download the Data Driven training set values from S3
The independent variables for the training set
```{r download-training-set-values}
# URL of dataset
data.url <-"https://s3.amazonaws.com/drivendata/data/7/public/4910797b-ee55-40a7-8668-10efd5c1b960.csv"

# File name of dataset
file.name <- "./data/src-training-set-values.csv"

# Download the file if it doesn't exist
if( !file.exists(file.name) ) {
    download(data.url, dest = file.name, mode = "wb")
}

```
\newline


## Download the Data Driven training set labels from S3
The dependent variable (status_group) for each of the rows in Training set values
```{r download-training-set-labels}
# URL of dataset
data.url <-"https://s3.amazonaws.com/drivendata/data/7/public/0bf8bc6e-30d0-4c50-956a-603fc693d966.csv"

# File name of dataset
file.name <- "./data/src-training-set-labels.csv"

# Download the file if it doesn't exist
if( !file.exists(file.name) ) {
    download(data.url, dest = file.name, mode = "wb")
}

```
\newline


## Download the Data Driven test set values from S3
The independent variables that need predictions
```{r download-test-set-values}
# URL of dataset
data.url <-"https://s3.amazonaws.com/drivendata/data/7/public/702ddfc5-68cd-4d1d-a0de-f5f566f76d91.csv"

# File name of dataset
file.name <- "./data/src-test-set-values.csv"

# Download the file if it doesn't exist
if( !file.exists(file.name) ) {
    download(data.url, dest = file.name, mode = "wb")
}

```
\newline


## Confirm files were downloaded
```{r confirm-download}
list.files("./data")

file.size("./data/src-training-set-values.csv")
file.size("./data/src-training-set-labels.csv")
file.size("./data/src-test-set-values.csv")
```
