---
title: 'MSDS-692: Data Science Practicum I'
subtitle: 'Exploratory Data Analysis'
author: "Thomas Skowronek"
date: "January 27, 2018"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
  word_document: default
---


```{r setup-layout, cache=FALSE, echo=FALSE}
# Define the report layout.
library(knitr)

# Set global chunk options: images sizes for html and docx
output <- opts_knit$get("rmarkdown.pandoc.to")

if (output=="html") opts_chunk$set(fig.width=10,  fig.height=6)
if (output=="docx") opts_chunk$set(fig.width=10,  fig.height=6)

# Set the printable width
options(width = 95)
```
\newline


# Setup
Set the working directory, clear all existing objects in the workspace and set the seed for reproducibility.  Lastly, load the necessary libraries.
```{r env-config, warning=FALSE, message=FALSE}
# Set the working directory
setwd("./")

# Clear all existing objects in the workspace
rm(list = ls())

# Set the seed for reproducible results
set.seed(1009)

# Load libraries
library(dplyr)
library(ggplot2)
library(ggmap)
library(scales)
library(RColorBrewer)
```
\newline


# Attribute Definitions:
* id - Well identifier
* amount_tsh - Total static head (amount water available to waterpoint)
* date_recorded - The date the row was entered
* funder - Who funded the well
* gps_height - Altitude of the well
* installer - Organization that installed the well
* longitude - GPS coordinate
* latitude - GPS coordinate
* wpt_name - Name of the waterpoint if there is one
* num_private -
* basin - Geographic water basin
* subvillage - Geographic location
* region - Geographic location
* region_code - Geographic location (coded)
* district_code - Geographic location (coded)
* lga - Geographic location
* ward - Geographic location
* population - Population around the well
* public_meeting - True/False
* recorded_by - Group entering this row of data
* scheme_management - Who operates the waterpoint
* scheme_name - Who operates the waterpoint
* permit - If the waterpoint is permitted
* construction_year - Year the waterpoint was constructed
* extraction_type - The kind of extraction the waterpoint uses
* extraction_type_group - The kind of extraction the waterpoint uses
* extraction_type_class - The kind of extraction the waterpoint uses
* management - How the waterpoint is managed
* management_group - How the waterpoint is managed
* payment - What the water costs
* payment_type - What the water costs
* water_quality - The quality of the water
* quality_group - The quality of the water
* quantity - The quantity of water
* quantity_group - The quantity of water
* source - The source of the water
* source_type - The source of the water
* source_class - The source of the water
* waterpoint_type - The kind of waterpoint
* waterpoint_type_group - The kind of waterpoint
* status_group - The operational status of the well
    + functional - the waterpoint is operational and there are no repairs needed
    + functional needs repair - the waterpoint is operational, but needs repairs
    + non functional - the waterpoint is not operational

\newline


# Load the Datasets
```{r load-data}
train.values <- read.csv("./data/clean-training-set-values.csv", header = TRUE, na.strings = "NA")
train.labels <- read.csv("./data/src-training-set-labels.csv", header = TRUE)
test.values <- read.csv("./data/clean-test-set-values.csv", header = TRUE, na.strings = "NA")
```
\newline


# Analysis
For the analysis, the training values and labels will be merged, and the exploration will be conducted on the combined dataset.
## Merge the training values and labels
```{r merge-datasets}
training.data <- merge(train.values, train.labels)
```
\newline


## View the Summary Statistics
The following variables contain missing values:

* funder - 4,414
* installer - 4,435
* wpt_name - 3,565
* subvillage - 371
* scheme_management - 3,877
* scheme_name - 28,166
* construction_year - 20,709

```{r summary-stats}
summary(training.data)
```
\newline


## Pump Status Distribution
An inspection of the pump status group shows that just over half of the water pumps are operational at 54.3%.
```{r pump-status}
ggplot(data=training.data, aes(x=status_group)) + 
    geom_bar(aes(y = (..count..)/sum(..count..)), fill=brewer.pal(3, "Set1"), alpha = 1/2) + 
    geom_text(aes(y = ((..count..)/sum(..count..)), label = scales::percent((..count..)/sum(..count..))), 
              stat = "count", vjust = -0.25) +
    scale_y_continuous(labels = percent) +
    theme(axis.text.y=element_blank(), axis.ticks=element_blank(), axis.title.y=element_blank()) +
    ggtitle("Pump Status Distribution") +
    theme(plot.title = element_text(hjust = 0.5, face="bold")) +
    xlab("Operational Status") + ylab("Percent")
```
\newline


## Inspect Categorical Variables
The next few plots compare categorical variables by pump status groups.

### Water Basins
Ploting the water pumps by basin and separating them by status shows that the Pangani basin has the most functional water pumps.  The Lake Victoria basin has the highest quantity of water pumps that require repair or are not functional.  In addition, Lake Rukwa, Lake Tanganyika, and Ruvuma / Southern Coast have more non-functional pumps than functional pumps.
```{r basin-plot}
ggplot(training.data, aes(x=basin, fill = basin)) + geom_bar(alpha = 1/2) + 
  scale_fill_brewer(palette="Set1") + facet_grid(~status_group) +
  theme(axis.text.x=element_text(angle = -45, hjust = 0)) +
  xlab("") + ylab("Pump Count") + guides(fill=guide_legend(title="Basin"))
```


## Visualize Pumps in Tanzania
Plot all of the pumps regardless of status group to determine if the are clustered in certain parts of the country.
```{r  map, warning=FALSE, fig.width=12, fig.height=8}
# Center of Tanzania
map <- get_map(c(lon = 34.888822, lat = -6.369028), source = "google", zoom = 6)

ggmap(map) + geom_point(aes(x = longitude, y = latitude, colour = status_group), data = training.data, alpha = .5)
```