---
title: 'Predicting the Operational Status of Tanzanian Water Pumps'
subtitle: 'Random Forest'
author: "Thomas Skowronek"
date: "February 09, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---


```{r setup-layout, cache=FALSE, echo=FALSE}
# Define the report layout.
library(knitr)

# Set global chunk options: images sizes for html and docx
output <- opts_knit$get("rmarkdown.pandoc.to")

if (output=="html") opts_chunk$set(fig.width=10,  fig.height=6)
if (output=="docx") opts_chunk$set(fig.width=10,  fig.height=6)

# Set the printable width
options(width = 95)
```
\newline


# Setup
Set the working directory, clear all existing objects in the workspace and set the seed for reproducibility.  Lastly, load the necessary libraries.
```{r env-config, warning=FALSE, message=FALSE}
# Set the working directory
setwd("./")

# Clear all existing objects in the workspace
rm(list = ls())

# Set the seed for reproducible results
set.seed(1009)

# Load libraries
library(dplyr)
library(randomForest)
library(caret)
```
\newline


# Load the Datasets
```{r load-data, cache=TRUE}
train.values <- read.csv("../data/clean-training-set-values.csv", header = TRUE, na.strings = "NA")
train.values$permit <- as.character(train.values$permit)
train.values$permit[is.na(train.values$permit)] <- "UNKNOWN"
train.values$permit <- as.factor(train.values$permit)
train.values$construction_year[is.na(train.values$construction_year)] <- 0
train.values$population[is.na(train.values$population)] <- 0
train.values$scheme_management <- as.character(train.values$scheme_management)
train.values$scheme_management[is.na(train.values$scheme_management)] <- "UNKNOWN"
train.values$scheme_management <- as.factor(train.values$scheme_management)


train.labels <- read.csv("../data/src-training-set-labels.csv", header = TRUE)

# Merge the data values and labels
training.data <- merge(train.values, train.labels)

# Drop the id attribute
training.data <- training.data %>% select(-one_of(c("id")))

str(training.data)
summary(training.data)
```
\newline


# Split the Training Data
Create training and testing subsets of the data.
```{r train-test}
# Create an 80/20 split
random.idx <- createDataPartition(training.data$status_group, p = .80, list = FALSE)
pump.train <- training.data[ random.idx, ]
pump.test <- training.data[-random.idx, ]

# Verify similar proportions of status_group for training vs.testing
prop.table(table(pump.train$status_group))
prop.table(table(pump.test$status_group))
```
\newline


# Random Forest Model #1
Create random forest model using a subset of attributes.
```{r model-1}
pump.model.1 <- randomForest(status_group ~ longitude + latitude + extraction_type_group + construction_year +
                                quality_group + quantity + waterpoint_type, 
                              data = pump.train, importance = TRUE, ntree = 10, nodesize = 2)

# Inspect the model
pump.model.1
#summary(pump.model.1)

# Examine the importance of the predictor attributes
importance(pump.model.1)
varImpPlot(pump.model.1)

# Use the model to create predictions against the test dataset
pump.predict.1 <- predict(pump.model.1, select(pump.test, -status_group))

# Evaluate the model using a confusion matrix
confusionMatrix(pump.predict.1, pump.test$status_group, mode = "everything")
```
\newline


# Random Forest Model #2
Create random forest model using a larger subset of attributes and trees.
```{r model-2}
pump.model.2 <- randomForest(status_group ~ basin + region + permit + extraction_type_class + quality_group + 
                               quantity + source + waterpoint_type + population + amount_tsh + gps_height + 
                               district_code + scheme_management + payment + longitude + latitude,
                              data = pump.train, importance = TRUE, ntree = 500)

# Inspect the model
pump.model.2
#summary(pump.model.2)

# Examine the importance of the predictor attributes
importance(pump.model.2)
varImpPlot(pump.model.2)

# Evaluate the model using a confusion matrix
pump.predict.2 <- predict(pump.model.2, select(pump.test, -status_group))

# Evaluate the model using a confusion matrix
confusionMatrix(pump.predict.2, pump.test$status_group, mode = "everything")
```


