---
title: 'Predicting the Operational Status of Tanzanian Water Pumps'
subtitle: 'XGBoost'
author: "Thomas Skowronek"
date: "February 12, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---


```{r setup-layout, cache=FALSE, echo=FALSE}
# Define the report layout.
library(knitr)

# Set global chunk options: images sizes for html and docx
output <- opts_knit$get("rmarkdown.pandoc.to")

if (output=="html") opts_chunk$set(fig.width=10,  fig.height=6)
if (output=="docx") opts_chunk$set(fig.width=10,  fig.height=6)

# Set the printable width
options(width = 95)
```
\newline


# Setup
Set the working directory, clear all existing objects in the workspace and set the seed for reproducibility.  Lastly, load the necessary libraries.
```{r env-config, warning=FALSE, message=FALSE}
# Set the working directory
setwd("./")

# Clear all existing objects in the workspace
rm(list = ls())

# Set the seed for reproducible results
set.seed(1009)

# Load libraries
library(dplyr)
library(Matrix)
library(xgboost)
library(caret)
```
\newline


# Load the Datasets
```{r load-data}
# Variable name: training.data
source("./modules/load-training-data.R")
str(training.data)
summary(training.data)

# Variable name: test.values
source("./modules/load-test-data.R")
str(test.values)
summary(test.values)
```
\newline


# Split the Training Data
Create training and testing subsets of the data.
```{r train-test}
# Create an 80/20 split
random.idx <- createDataPartition(training.data$status_group, p = .80, list = FALSE)

pump.train <- training.data[ random.idx, ]
pump.test <- training.data[-random.idx, ]

# Verify similar proportions of status_group for training vs.testing
prop.table(table(pump.train$status_group))
prop.table(table(pump.test$status_group))
```
\newline


# Convert datasets to numeric vectors
XGboost requires numberic vectors.
```{r data-conversion}
predictors <- c("longitude", "latitude", "construction_year", "extraction_type_group",
                "quality_group", "quantity", "waterpoint_type", "status_group")

# Training data
train.subset <- pump.train %>% select(predictors)
train.sparse.matrix <- sparse.model.matrix(status_group ~ .-1, data = train.subset)
train.output.vector <- ifelse(train.subset$status_group == "functional", 0, 
                        ifelse(train.subset$status_group == "non functional", 1, 2))

# Testing data
test.subset <- pump.test %>% select(predictors)
test.sparse.matrix <- sparse.model.matrix(status_group ~ .-1, data = test.subset)
test.output.vector <- ifelse(test.subset$status_group == "functional", 0, 
                        ifelse(test.subset$status_group == "non functional", 1, 2))
```
\newline


# XGBoost Model #1
Create xgboost model using a subset of attributes.
```{r model-1}

pump.model.1 <- xgboost(data = train.sparse.matrix, label = train.output.vector, max_depth = 15, eta = 1, 
               nthread = 2, nrounds = 25, num_class = 3, objective = "multi:softmax")

# Examine the importance of the predictor attributes
importance <- xgb.importance(feature_names = colnames(train.sparse.matrix), model = pump.model.1); importance

# Use the model to create predictions against the test dataset
pump.predict.1 <- predict(pump.model.1, newdata = test.sparse.matrix)

# Evaluate the model using a confusion matrix
confusionMatrix(pump.predict.1, test.output.vector, mode = "everything")
```
\newline


# Full set of us usable predictors
Used by the next two models
```{r full-predictors}
# "ward",
predictors <- c("amount_tsh", "funder.bin", "gps_height", "installer.bin", "longitude", "latitude",
                "basin", "region", "region_code", "district_code", "lga", "population", 
                "public_meeting", "scheme_management", "permit", "construction_year", "extraction_type", 
                "extraction_type_group", "extraction_type_class", "management", "management_group", 
                "payment", "payment_type", "water_quality", "quality_group", "quantity", 
                "quantity_group", "source", "source_type", "source_class", "waterpoint_type", 
                "waterpoint_type_group", "status_group")

train.subset <- pump.train %>% select(predictors)
train.sparse.matrix <- sparse.model.matrix(status_group ~ .-1, data = train.subset)
train.output.vector <- ifelse(train.subset$status_group == "functional", 0, 
                        ifelse(train.subset$status_group == "non functional", 1, 2))

test.subset <- pump.test %>% select(predictors)
test.sparse.matrix <- sparse.model.matrix(status_group ~ .-1, data = test.subset)
test.output.vector <- ifelse(test.subset$status_group == "functional", 0, 
                        ifelse(test.subset$status_group == "non functional", 1, 2))
```
\newline


# XGBoost Model #2
Create xgboost model using a full set attributes.
```{r model-2}
ptm.start <- proc.time()

pump.model.2 <- xgboost(data = train.sparse.matrix, label = train.output.vector, max_depth = 15, eta = 0.3,
                        nrounds = 100, num_class = 3, objective = "multi:softmax")

# Examine the importance of the predictor attributes
importance <- xgb.importance(feature_names = colnames(train.sparse.matrix), model = pump.model.2); importance

# Use the model to create predictions against the test dataset
pump.predict.2 <- predict(pump.model.2, newdata = test.sparse.matrix)

# Evaluate the model using a confusion matrix
confusionMatrix(pump.predict.2, test.output.vector, mode = "everything")

ptm.end <- proc.time() - ptm.start
ptm.end[3] / 60
```
\newline