---
title: 'Predicting the Operational Status of Tanzanian Water Pumps'
subtitle: 'Support Vector Machines'
author: "Thomas Skowronek"
date: "February 17, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---


```{r setup-layout, cache=FALSE, echo=FALSE}
# Define the report layout.
library(knitr)

# Set global chunk options: images sizes for html and docx
output <- opts_knit$get("rmarkdown.pandoc.to")

if (output=="html") opts_chunk$set(fig.width=10,  fig.height=6)
if (output=="docx") opts_chunk$set(fig.width=10,  fig.height=6)

# Set the printable width
options(width = 95)
```
\newline


# Setup
Set the working directory, clear all existing objects in the workspace and set the seed for reproducibility.  Lastly, load the necessary libraries.
```{r env-config, warning=FALSE, message=FALSE}
# Set the working directory
setwd("./")

# Clear all existing objects in the workspace
rm(list = ls())

# Set the seed for reproducible results
set.seed(1009)

# Load libraries
library(dplyr)
library(e1071)
library(caret)
```
\newline


# Load the Datasets
```{r load-data}
# Variable name: training.data
source("./modules/load-training-data.R")
str(training.data)
summary(training.data)

# Variable name: test.values
source("./modules/load-test-data.R")
str(test.values)
summary(test.values)
```
\newline


# Split the Training Data
Create training and testing subsets of the data.
```{r split-data}
# Create an 80/20 split
random.idx <- createDataPartition(training.data$status_group, p = .80, list = FALSE)
pump.train <- training.data[ random.idx, ]
pump.test <- training.data[-random.idx, ]

# Verify similar proportions of status_group for training vs.testing
prop.table(table(pump.train$status_group))
prop.table(table(pump.test$status_group))

predictors <- c("longitude", "latitude", "construction_year", "extraction_type_group",
                "quality_group", "quantity", "waterpoint_type", "status_group")

# Subsetting the attributes because attributes that are not used and contain NA affect predictions
train.subset <- pump.train %>% select(predictors)
test.subset <- pump.test %>% select(predictors)
```
\newline


# SVM Model #1
Create SVM model using a subset of attributes.
```{r model-1}
ptm.start <- proc.time()

pump.model.1 <- svm(status_group ~ longitude + latitude + extraction_type_group + construction_year + 
                      quality_group + quantity + waterpoint_type, data=train.subset)

# Inspect the model
summary(pump.model.1)

# Use the model to create predictions against the test dataset
pump.predict.1 <- predict(pump.model.1, newdata = select(test.subset, -status_group))
summary(pump.predict.1)

#pump.predict.1[1] <- "functional needs repair"

# Evaluate the model using a confusion matrix
confusionMatrix(table(pump.predict.1, pump.test$status_group), mode = "everything")

ptm.end <- proc.time() - ptm.start
ptm.end[3] / 60
```
\newline


# Full set of us usable predictors
Used by the next two models
```{r full-predictors}
# "ward",
predictors <- c("amount_tsh", "funder.bin", "gps_height", "installer.bin", "longitude", "latitude",
                "basin", "region", "region_code", "district_code", "lga",  "population", 
                "public_meeting", "scheme_management", "permit", "construction_year", "extraction_type", 
                "extraction_type_group", "extraction_type_class", "management", "management_group", 
                "payment", "payment_type", "water_quality", "quality_group", "quantity", 
                "quantity_group", "source", "source_type", "source_class", "waterpoint_type", 
                "waterpoint_type_group", "status_group")

train.subset <- pump.train %>% select(predictors)
test.subset <- pump.test %>% select(predictors)
```
\newline


# SVM Model #2
Create SVM model using the full set of attributes with a radial kernal.
```{r model-2}
ptm.start <- proc.time()

pump.model.2 <- svm(status_group ~ ., data=train.subset, kernel = "radial")

# Inspect the model
summary(pump.model.2)

# Use the model to create predictions against the test dataset
pump.predict.2 <- predict(pump.model.2, newdata = select(test.subset, -status_group))
summary(pump.predict.2)

#pump.predict.2[1] <- "functional needs repair"

# Evaluate the model using a confusion matrix
confusionMatrix(table(pump.predict.2, pump.test$status_group), mode = "everything")

ptm.end <- proc.time() - ptm.start
ptm.end[3] / 60
```
\newline


# SVM Model #3
Create SVM model using the full set of attributes with a polynomial kernal.
```{r model-3}
ptm.start <- proc.time()

pump.model.3 <- svm(status_group ~ ., data=train.subset, kernel = "polynomial")

# Inspect the model
summary(pump.model.3)

# Use the model to create predictions against the test dataset
pump.predict.3 <- predict(pump.model.3, newdata = select(test.subset, -status_group))
summary(pump.predict.3)

#pump.predict.3[1] <- "functional needs repair"

# Evaluate the model using a confusion matrix
confusionMatrix(table(pump.predict.3, pump.test$status_group), mode = "everything")

ptm.end <- proc.time() - ptm.start
ptm.end[3] / 60
```
\newline


# SVM Model #4
Create SVM model using the full set of attributes with a sigmoid kernal.
```{r model-4}
ptm.start <- proc.time()

pump.model.4 <- svm(status_group ~ ., data=train.subset, kernel = "sigmoid")

# Inspect the model
summary(pump.model.4)

# Use the model to create predictions against the test dataset
pump.predict.4 <- predict(pump.model.4, newdata = select(test.subset, -status_group))
summary(pump.predict.4)

#pump.predict.4[1] <- "functional needs repair"

# Evaluate the model using a confusion matrix
confusionMatrix(table(pump.predict.4, pump.test$status_group), mode = "everything")

ptm.end <- proc.time() - ptm.start
ptm.end[3] / 60
```
\newline


# SVM Model #5
Create SVM model using the full set of attributes with 10-fold cross validation.
```{r model-5}
ptm.start <- proc.time()

pump.model.5  <- train(status_group ~ ., data = train.subset, method = "svmRadial", 
                       trControl = trainControl(method = "cv", number = 10),
                       tuneLength = 2)

# Inspect the model
summary(pump.model.5)

# Examine the importance of the predictor attributes
varImp(pump.model.5, scale = FALSE)
plot(pump.model.5, top = 20)

# Use the model to create predictions against the test dataset
pump.predict.5 <- predict(pump.model.5, newdata = select(test.subset, -status_group))
summary(pump.predict.5)

#pump.predict.5[1] <- "functional needs repair"

# Evaluate the model using a confusion matrix
confusionMatrix(table(pump.predict.5, pump.test$status_group), mode = "everything")

# Use the model to create predictions against the real test dataset
pump.predict.test.5 <- predict(pump.model.5, select(test.values, -id))

# Create submission for Data Driven
submission.5 <- data.frame(id = test.values$id, status_group = pump.predict.test.5)
write.csv(submission.5, file = "../data/svm-model-5.csv", row.names=FALSE)

ptm.end <- proc.time() - ptm.start
ptm.end[3] / 60
```
\newline
