---
title: 'Predicting the Operational Status of Tanzanian Water Pumps'
subtitle: 'H2O.ai Deep Learning'
author: "Thomas Skowronek"
date: "February 18, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---


```{r setup-layout, cache=FALSE, echo=FALSE}
# Define the report layout.
library(knitr)

# Set global chunk options: images sizes for html and docx
output <- opts_knit$get("rmarkdown.pandoc.to")

if (output=="html") opts_chunk$set(fig.width=10,  fig.height=6)
if (output=="docx") opts_chunk$set(fig.width=10,  fig.height=6)

# Set the printable width
options(width = 95)
```
\newline


# Setup
Set the working directory, clear all existing objects in the workspace and set the seed for reproducibility.  Lastly, load the necessary libraries.
```{r env-config, warning=FALSE, message=FALSE}
# Set the working directory
setwd("./")

# Clear all existing objects in the workspace
rm(list = ls())

# Set the seed for reproducible results
set.seed(1009)

# Load libraries
library(dplyr)
library(h2o)
library(caret)
library(ROCR)
```
\newline


# Load the Datasets
```{r load-data, cache=TRUE}
# Variable name: training.data
source("./modules/load-training-data.R")
str(training.data)
summary(training.data)

# Variable name: test.values
source("./modules/load-test-data.R")
str(test.values)
summary(test.values)
```
\newline


# H2O setup
```{r h20}
# Disable progress bars for Rmd
h2o.no_progress()  

# Initialize and connect to H2O
local.h20 <- h2o.init()


# predictors <- c("longitude", "latitude", "construction_year", "extraction_type_group",
#                 "quality_group", "quantity", "waterpoint_type", "status_group")

predictors <- c("amount_tsh", "funder.bin", "gps_height", "installer.bin", "longitude", "latitude",
                "basin", "region", "region_code", "district_code", "lga",  "population", 
                "public_meeting", "scheme_management", "permit", "construction_year", "extraction_type", 
                "extraction_type_group", "extraction_type_class", "management", "management_group", 
                "payment", "payment_type", "water_quality", "quality_group", "quantity", 
                "quantity_group", "source", "source_type", "source_class", "waterpoint_type", 
                "waterpoint_type_group", "status_group")

train.subset <- training.data %>% select(predictors)

# Create H2OFrame
pump.hex <- as.h2o(train.subset, destination_frame="pump.hex")

# Create a 80/20 test/train datasets
train.test <- h2o.splitFrame(data = pump.hex, ratios = 0.80, seed = 1009)
pump.train <- train.test[[1]]
pump.test <- train.test[[2]]
pump.test.df <- as.data.frame(pump.test)

# Identify the response and predictor variables
y <- "status_group"
x <- setdiff(names(train.subset), c(y, "class"))
print(x)
```
\newline


# Model 1
Two hundred hidden layers with 200 nodes each
```{r model-1}
ptm.start <- proc.time()

# Create the DL model
pump.fit1 <- h2o.deeplearning(x = x, y = y, training_frame = pump.train, 
                                model_id = "pump.fit1", hidden = c(200,200,200),
                                seed = 1009)

# Obtain the predictions from the fitted model
pump.pred1 <- h2o.predict(pump.fit1, pump.test)
pump.pred1.df <- as.data.frame(pump.pred1)
pump.pred1

# Show the confustion matrix with additional performance metrics
confusionMatrix(pump.pred1.df$predict, pump.test.df$status_group,  mode = "everything")

ptm.end <- proc.time() - ptm.start
ptm.end[3] / 60
```
\newline


# Model 2
Two hundred hidden layers with 200 nodes each and early stopping disabled
```{r model-2}
ptm.start <- proc.time()

# Create the DL model
pump.fit2 <- h2o.deeplearning(x = x, y = y, training_frame = pump.train, 
                                model_id = "pump.fit2", hidden = c(200,200,200),
                                epochs = 50, stopping_rounds = 0, seed = 1009)

# Obtain the predictions from the fitted model
pump.pred2 <- h2o.predict(pump.fit2, pump.test)
pump.pred2.df <- as.data.frame(pump.pred2)
pump.pred2

# Show the confustion matrix with additional performance metrics
confusionMatrix(pump.pred2.df$predict, pump.test.df$status_group,  mode = "everything")

ptm.end <- proc.time() - ptm.start
ptm.end[3] / 60
```
\newline


# Model 3
Two hundred hidden layers with 200 nodes each, early stopping enabled with defined stopping criteria
```{r model-3}
ptm.start <- proc.time()

# Create the DL model
pump.fit3 <- h2o.deeplearning(x = x, y = y, training_frame = pump.train, 
                                model_id = "pump.fit3", hidden = c(200,200,200),
                                epochs = 50, seed = 1009, 
                                nfolds = 3,                            #used for early stopping
                                score_interval = 1,                    #used for early stopping
                                stopping_rounds = 5,                   #used for early stopping
                                stopping_metric = "misclassification", #used for early stopping
                                stopping_tolerance = 1e-3)             #used for early stopping

# Obtain the predictions from the fitted model
pump.pred3 <- h2o.predict(pump.fit3, pump.test)
pump.pred3.df <- as.data.frame(pump.pred3)
pump.pred3

# Show the confustion matrix with additional performance metrics
confusionMatrix(pump.pred3.df$predict, pump.test.df$status_group,  mode = "everything")

ptm.end <- proc.time() - ptm.start
ptm.end[3] / 60
```
\newline



# Shutdown H2O
```{r shutdown}
# Shutdown the cluster
h2o.shutdown(prompt = FALSE)
```


# References
LeDell, E. (2017). H2O Deep Learning. Retrieved from https://github.com/ledell/sldm4-h2o/blob/master/sldm4-deeplearning-h2o.Rmd